---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# InterpretMSSpectrum

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/InterpretMSSpectrum)](https://CRAN.R-project.org/package=InterpretMSSpectrum)
[![R-CMD-check](https://github.com/janlisec/InterpretMSSpectrum/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/janlisec/InterpretMSSpectrum/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/janlisec/InterpretMSSpectrum/branch/main/graph/badge.svg?token=NSY6DITZVH)](https://codecov.io/gh/janlisec/InterpretMSSpectrum)
<!-- badges: end -->

The goal of **InterpretMSSpectrum** is to provides a set of R functions to annotate
mass spectra from Electrospray-Ionization and Atmospheric-Pressure-Chemical-Ionization 
derived data in positive and negative ionization mode.

## Installation

You can install the [CRAN](https://cran.r-project.org/package=InterpretMSSpectrum) 
version of **InterpretMSSpectrum** using:
```
install.packages("InterpretMSSpectrum")
```
or the development version from [GitHub](https://github.com/) using:
```
# devtools is required to install from GitHub
# install.packages("devtools")
devtools::install_github("janlisec/InterpretMSSpectrum")
```

## Example

In the simplest case **InterpretMSSpectrum** will provide an informed guess
for the potential sum formula of an arbitrary mass spectrum.

```{r exmpl1}
# load APCI test data
apci_spectrum <- InterpretMSSpectrum::apci_spectrum

# find the most probable sum formula for the spectrum
# (will print to the console and open a new plot)
InterpretMSSpectrum::InterpretMSSpectrum(spec=apci_spectrum)
```

The function can be tweaked with numerous parameters to limit the results, speed
up calculations and more. 

The other high level function of the package allows to predict the precursor of 
ESI spectra.

```{r exmpl2}
# load ESI test data
esi_spectrum <- InterpretMSSpectrum::esi_spectrum

# find the most likely precursor for the spectrum
(fmr <- InterpretMSSpectrum::findMAIN(spec=esi_spectrum))
plot(fmr)
```

Also `findMAIN` provides multiple user options to use individual adduct lists, 
thresholds and rule sets.

Finally, `InterpretMSSpectrum` provides a number of helper functions, some of 
which are listed below.

```{r exmpl3}
# to count the chemical elements within a character vector of sum formulas
InterpretMSSpectrum::CountChemicalElements(x = "C6H12O6")
sapply(c("C6H12O6", "CH3Cl"), InterpretMSSpectrum::CountChemicalElements, ele=c("C","H","O"))

# to get the exact mass for a sum formula
InterpretMSSpectrum::get_exactmass(c("C6H12O6", "Na", "H1"))

# to check if one formula is contained in another (i.e. as fragment)
InterpretMSSpectrum:::is.subformula("H2O", "HCOOH")
```

## Detailed documentation

You might read the publications on either
[APCI spectra processing](https://doi.org/10.1021/acs.analchem.6b02743) which
explains the idea of using the in source fragments for prediction of potential 
sum formulas or on
[ESI spectra processing](https://doi.org/10.1002/rcm.7905) which 
explains the strategy to infer the correct precursor of ESI mass spectra.
